---
title: "An example exploratory analysis script"
author: "Andreas Handel"
date: "10/28/2021"
output: html_document
---

  
  

This script loads the cleaned data and does some exploring.

Note that at times it is better to combine the cleaning and exploring code into a single R or Rmd file. Or split things in any other logical way.

You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it.

As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. Figures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data.

Start by exploring one variablea at a time. Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables. 

Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(ggplot2)
```


Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```





# Data exploration through tables

Showing a bit of code to produce and save a summary table.

This uses a bit of code trickery to get things into the shape of a data frame (for easier saving/showing in manuscript)

```{r}
summary_df = data.frame(do.call(cbind, lapply(mydata, summary)))
summarytable_file = here("results", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```



# Data exploration through figures

Histogram plots for the continuous outcomes. 

Height first.

```{r}
p1 <- mydata %>% ggplot(aes(x=Height)) + geom_histogram() 
plot(p1)
figure_file = here("results","height_distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

Now weights.

```{r}
p2 <- mydata %>% ggplot(aes(x=Weight)) + geom_histogram() 
plot(p2)
figure_file = here("results","weight_distribution.png")
ggsave(filename = figure_file, plot=p2) 
```

Now height as function of weight.

```{r}
p3 <- mydata %>% ggplot(aes(x=Height, y=Weight)) + geom_point() + geom_smooth(method='lm')
plot(p3)
figure_file = here("results","height_weight.png")
ggsave(filename = figure_file, plot=p3) 
```

Once more height as function of weight, stratified by sex.

```{r}
p4 <- mydata %>% ggplot(aes(x=Height, y=Weight, color = Sex)) + geom_point() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results","height_weight_stratified.png")
ggsave(filename = figure_file, plot=p4) 
```



